### 2021-07-05 doc/thesis/chapters/dsp.tex
superhet and zero-if

\subsection{Direct-Conversion Receiver}

@@

\subsection{Superheterodyne Receiver}

The superheterodyne receiver is often found in the literature with short forms like \textit{superhet}, or simply \textit{super}.
The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\

The IF is usually set to a fixed frequency.
This results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be flexible in frequency.

Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
The antenna signal first runs through an RF amplifier and filter.
It selects the desired frequency band and filters out any unwanted signals outside of this band.
The next step is the mixer, which is fed by a local oscillator.
This shifts, or down-converts, the antenna signal down to the IF.
Afterwards, the IF signal is amplified and filtered again.
This is to remove the side-products of the mixer and may also be called as image rejection filter.
The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.
In the shown case, which may be an FM broadcast receiver, an audio output is generated.

\includepicture [1.0] [0] {Block design of a superhetorodyne receiver.} {bd_superhet} {img/superheterodyne-receiver-block-diagram.png}
%https://www.electronics-notes.com/articles/radio/superheterodyne-receiver/block-diagram.php

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Superheterodyne Receiver}

The superheterodyne receiver is often referred to as \textit{superhet}, or \textit{super} in the literature.
The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\

The IF is usually set to a fixed frequency, which results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be adapted in frequency.
Therefore, a superhet receiver is able to select narrow band signals, even if the bands' surrounding frequencies are being used by other signals.
This is one of the superhets' main advantages.
A disadvantage is that the signal noise around the image frequency band is translated into the IF output band.
Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.

Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
The antenna signal first runs through a band selection filter.
It selects the desired frequency band and filters out any unwanted signals outside of this band.
A low-noise amplifier (LNA) may be used to boost the signal strenght.
The next step is the mixer, which is fed by a local oscillator (LO).
This shifts, or down-converts, the antenna signal down to the IF.
Afterwards, the IF signal is amplified and filtered again to remove the side-products of the mixer, namely the image replicas.
Thus, it is also called an image rejection filter.
The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved \cite{PassosFábio2020AHSo}.\\

The depicted block diagram may represent an FM broadcast receiver, where an audio output is generated from the antenna input.

\includepicture [1.0] [0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/superheterodyne-receiver-block-diagram-2.png}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Direct-Conversion Receiver}

@@

The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called $\frac{1}{f}$-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
These effects are not explained into more detail here, since this is not the focus of this thesis.

\includepicture [0.9] [0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/direct-conversion-receiver-block-diagram.png}
